Validating Complex Interval-Based Temporal Queries over Streams

### 2013-05-10 paper/paper.tex
Final draft

\begin{document}
\maketitle
\begin{abstract}
%Abstract here.
Streaming databases have been growing in popularity as the availability of large datasets gathered in real-time has steadily increased.  Along with the ability to gather and store this data, one must also consider the various concerns involved in processing this data.  More specifically, the performance of queries over such databases is highly dependent on the complexity of the query and the database itself.  These considerations are common in analyses of database systems.  However, on 
\end{abstract}
\section{Introduction}
The ability to query a database for temporal events has become increasingly important as the availability of streaming data has skyrocketed across multiple domains. While event stream processing has seen much work~\cite{Agrawal:2008, Akdere:2008, Ding:2008, Wu:2006, Brenna:2007, Li:2009, Li:2010, Liu:2009}, it has mainly focused on the relations between instantaneous events. However, the modelling of events as single points in time is not always appropriate. Take, for example, a database which tracks a real-time, multi-processor system. In this system it is important to know when processors are and are not available for use, a task which is traditionally modelled as a series of constrained intervals. Thus, a database paradigm that can handle intervals as well as instantaneous events is necessary.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Streaming databases have been growing in popularity as the availability of large datasets gathered in real-time has steadily increased.  Along with the ability to gather and store this data, one must also consider the various concerns involved in processing this data.  More specifically, the performance of queries over such databases is highly dependent on the complexity of the query and the database itself.  These considerations are common in analyses of database systems.  However, one portion of query processing that seems to be often overlooked is query validation as it is generally assumed that a given query is valid.

With this in mind, this project seeks to discover whether or not the resources necessary to validate a query make it an entirely cost-prohibitive action, or, if not, whether an intelligent decision could be made on a case-by-case basis whether or not validation should occur at all.  In particular cases, e.g., it may still be faster to blindly run an invalid query on a small database.

This paper discusses various methods for validating temporal queries, presents an implementation of a temporal database with range queries, and describes an experiment to determine the time costs of adding a validation step to query plans.  Moreover, it can be seen that there does appear to be a point at which the size of the database is large enough to warrant query validation.
\end{abstract}
\section{Introduction}
The ability to query a database for temporal events has become increasingly important as the availability of streaming data has skyrocketed across multiple domains. While event stream processing has seen much work~\cite{Agrawal:2008, Akdere:2008, Ding:2008, Wu:2006, Brenna:2007, Li:2009, Li:2010, Liu:2009}, it has mainly focused on the relations between instantaneous events. However, the modelling of events as single points in time is not always appropriate. Take, for example, a database which tracks a real-time, multi-processor system. In this system it is important to know when processors are and are not available for use, a task which is traditionally modelled as a series of constrained intervals. Thus, a database paradigm that can handle intervals as well as instantaneous events is necessary.
